A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-01-21 paper/paper.tex
moving the observation of indexed = parametricity to section "our construction"

We observe that the process of construction of the type of $X_1$ from that of $X_0$, and from the type of $X_2$ to that of $X_1$, is similar to applying a binary parametricity translation and expecting the resulting translation to be inhabited. Indeed, parametricity interprets a type $A$ by a relation $A_\kstar$ over this type, and a term $t: A$ as an evidence that $A_\kstar(t)(t)$ is inhabited. In particular, $\U$ is interpreted as the relation ${\U}_\kstar$, which takes two types $A_L$ and $A_R$ and returns the type $A_L \rightarrow A_R \rightarrow \U$, while, for $A$ interpreted by $A_\kstar$ and $B$ interpreted by $B_\kstar$, a dependent function type $\Pi a: A.\, B$ is interpreted as the relation $(\Pi a: A.\, B)_\kstar$ that takes two functions $f_L$ and $f_R$ of type $\Pi a: A.\, B$, and expresses that these functions map related arguments in $A$ to related arguments in $B$:

@@

\begin{equation*}
  \begin{array}{llllr}
    X_0 & \defeq & X                & : & \U                                               \\
    X_1 & \defeq & X_\kstar         & : & {\U}_\kstar(X)(X)                                \\
    X_2 & \defeq & X_{\kstar\kstar} & : & ({\U}_\kstar(X)(X))_{\kstar}(X_\kstar)(X_\kstar) \\
    \ldots
  \end{array}
\end{equation*}
This tells us how the previous (informal) type given to $X_2$ could be rephrased so that it comes as the instance of a general recipe characterizing the type of all $X_i$.

>>>>

\subsection{Relating the indexed representation to parametricity}

We observe that the process of construction of the type of $X_1$ from that of $X_0$, and from the type of $X_2$ to that of $X_1$ in the last section, is similar to applying a binary parametricity translation and expecting the resulting translation to be inhabited. Indeed, parametricity interprets a type $A$ by a relation $A_\kstar$ over this type, and a term $t: A$ as an evidence that $A_\kstar(t)(t)$ is inhabited. In particular, $\U$ is interpreted as the relation ${\U}_\kstar$, which takes two types $A_L$ and $A_R$ and returns the type $A_L \rightarrow A_R \rightarrow \U$, while, for $A$ interpreted by $A_\kstar$ and $B$ interpreted by $B_\kstar$, a dependent function type $\Pi a: A.\, B$ is interpreted as the relation $(\Pi a: A.\, B)_\kstar$ that takes two functions $f_L$ and $f_R$ of type $\Pi a: A.\, B$, and expresses that these functions map related arguments in $A$ to related arguments in $B$:

@@

\begin{equation*}
  \begin{array}{llllr}
    X_0 & \defeq & X                & : & \U                                               \\
    X_1 & \defeq & X_\kstar         & : & {\U}_\kstar(X)(X)                                \\
    X_2 & \defeq & X_{\kstar\kstar} & : & ({\U}_\kstar(X)(X))_{\kstar}(X_\kstar)(X_\kstar) \\
    \ldots
  \end{array}
\end{equation*}
This tells us how the informal type given to $X_2$ in the previous section could be rephrased so that it comes as the instance of a general recipe characterizing the type of all $X_i$.
